\documentclass{article}
\usepackage[utf8]{inputenc}

\title{Project draft}
\author{Yue Wu}
\date{04 / 11 / 2018}

\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\maketitle

\section{Introduction}

The project targets on the effects of video games global sales. 

QR1: What companies hold more shares of market, what is the different between those companies from others.

QR2: What type of game is most popular. Why people like such type games.

QR3: How does the game market develop during the time changing. What is the effects of technology improving to game market.

QR4: How does game quality affects global sales.

\section{Literature Review}

"The history of video game development." This articles mentioned "game market crush" in 1980s, through the analysis of  the reason of the crush, we can study about how video game's quality affect market. 
There are some articles talk about the violence of video games. Normally violence only occur in certain types of games such as action and shooter games. We can learn that why violence games are more popular. 

\section{Data}

The data from Kaggle. variables include : Name, Platform, Year, Genre, Publisher, Sales, scores.
I consider the quality of games represented by scores.

\section{Empirical Methods}

Sales = b0 * platform + b1 * year + b3 * genre + b4 * publisher + b5 * scores

\section{Research Findings}

Game market grows fast in past 30 years.

Players preference of games also changes.

(More coming.....)

\section{Conclusion}

Quality(scores in Data) is the most effective term to global sales. 

(more coming.....)

\end{document}
